---
title: "Importance Sampling/ Impossible Sampling"
author: "MLi"
date: "`r format(Sys.time(), '%H:%M %d %B %Y')`"
output: html_document
---

# Introduction

The basic idea of importance sampling is to draw a large number of simulations from the approximation distribution (usually a MVN or t), decide how close each is to the target posterior distribution, and keep those close with higher probablility than for those farther away via the _importance ratio_ or _weights_.
Let X be a vector of parameters from a MVN(__X__, __V__) draw, the _importance ratio_ is the following:

$$\frac{L(X | data)}{\text{MVN}( X | \bf{X}, \bf{V})}$$

## Gaussian example

First simulation data

```{r sim data}

n <- 1e3
set.seed(112)
dat <- rnorm(n)

```

Fit the data via mle 

```{r mle,warning=FALSE}
library(bbmle)

nll <- function(m=0,s=1){
  return(-sum(dnorm(dat,m,s,log=TRUE)))
}
  
mlefit <- mle2(nll)

print(coef(mlefit))
print(vcov(mlefit))
```

The variances are so small, it can't even sample the space properly.
Following the book and use the crazy pseudovariance.

```{r impSamp}
samp_size <- 1e3

hh <- mlefit@details$hessian
V <- bdsmatrix:::gchol(MASS:::ginv(hh))
tV <- matrix(V@.Data,nrow=2)
V <- t(tV)

vv <- tV %*% V
mv_samps0 <- MASS:::mvrnorm(samp_size, mu = coef(mlefit), Sigma=vv)

print(summary(mv_samps0))


vv <- vcov(mlefit)
mv_samps <- MASS:::mvrnorm(samp_size, mu = coef(mlefit), Sigma=vv)

print(summary(mv_samps))
```

Even without the crazy pseudovariance method, we still can't simulate a good sample.

```{r imp_wts}
Posterior_negLL <- sapply(1:samp_size,function(x){nll(mv_samps[x,])})
approx_LL <- sapply(1:samp_size
  , function(x){
    mvtnorm::dmvnorm(mv_samps[x,]
      , mean = coef(mlefit)
      , sigma=vv
      , log = TRUE
      )
    }
)

Log_imp_wts <- -Posterior_negLL - approx_LL
print(summary(Log_imp_wts))
Log_scaled_imp_wts <- Log_imp_wts - max(Log_imp_wts)
print(summary(Log_scaled_imp_wts))

imp_wts <- exp(Log_scaled_imp_wts)
imp_wts <- imp_wts/sum(imp_wts) 
print(summary(imp_wts))

```

## Diagnostics

We can look at effective sample size and weighted quantiles.

```{r eff_samp}
eff_samp <- 1/sum(imp_wts^2)
print(eff_samp) 
```

Uh-oh!

```{r wq}
wq <- sapply(1:2
  , function(x){Hmisc::wtd.quantile(mv_samps[,x]
    , weights = imp_wts
    , probs = c(0.025, 0.975)
    , normwt = TRUE
    )
  }
)
print(t(wq))
print(confint(mlefit))

```

0 and 1 are not in the weighted quantiles and it looks nothing like the profile CIs.




