---
title: "Modelling Panther movements using moveHMM"
author: Michael Li
date: "`r format(Sys.time(), '%H:%M %d %B %Y')`"
---

# Intro / Testing moveHMM

moveHMM is a lot faster than depmixS4. I am going to try to replicate the BIC plots via moveHMM. 

```{r packages,warning=FALSE,message=FALSE}
t1 <- proc.time()
library(moveHMM)
library(dplyr)
library(ggplot2)
dat <- read.csv("ArchivePantherData.csv")

timenames <- levels(dat$Time)
dat$Time <- factor(dat$Time,levels=timenames,labels=0:23)
```

moveHMM requires a _moveData_ object for data frame. It takes the raw coordinates and computate the step length and turning angle using the _prepData_ function. Note: we can't simply feed out dataset in the fitting function, therefore, I class hack our dataframe. 
```{r dataframe, wearning=FALSE,message=FALSE}
class(dat) <- c("moveData","data.frame")
```
Lets keep it consistent with our workflow and use one cat at a time. Note, we can feed all the cats in at once, but let's save that for later.

```{r subseting}
cat1 <- dat %>% filter(animal_id==1) %>% transmute(time=Time,
                                                    step=Steplength.m.,
                                                    angle=NA)
## err 
class(cat1) <- c("moveData","data.frame")
```

Unlike depmixS4, we have to provide the starting parameters of each state. Let's keep it simple and start with two states. We can't fit univariate step-length HMMs, we can only fit multivariate step-length and turning angle HMMs. I can come up with three simple solution around this:

- change the turning angles to NA
- change the turning angles to 0 and give a very narrow VM distribution (ie. VM(0,1000)) for all states 
- just ignore it and hope it will not _bias_ the results

I like the first and second option the most. 
```{r parameters}
mu0 <- c(0,1)
sd0 <- c(0.1,0.5)
z0 <- c(0.1,0.2) ## not sure
step0 <- c(mu0,sd0,z0)

angleMean0 <- c(0,0)
kappa0 <- c(10000,10000)

angle0 <- c(angleMean0,kappa0)

hmm2s <- fitHMM(data=cat1,nbStates=2,stepPar0=step0,anglePar0=NULL,
                stepDist = "lnorm", angleDist = "none",
                formula = ~ 1)
                
hmm2s

## log-likelihood
hmm2s$mod$minimum

```


# Panther BIC results

```{r functions,echo=FALSE,warning=FALSE}

hmm <- function(n,dd){
  mu0 <- c(0:(n-1))
  sd0 <- rep(0.2,n)
  z0 <- rep(0.1,n)
  step0 <- c(mu0,sd0,z0)
  angle0 <- rep(c(0,10000),each=n)
  mod <- fitHMM(data=dd,nbStates=n,stepPar0=step0,anglePar0=NULL,
                stepDist = "lnorm", angleDist = "none",
                formula = ~ 1)
  return(mod)
}

hmm2s <- hmm(2,cat1)
hmm3s <- hmm(3,cat1)
hmm4s <- hmm(4,cat1)
hmm5s <- hmm(5,cat1)
hmm6s <- hmm(6,cat1)
hmm7s <- hmm(7,cat1)
hmm8s <- hmm(8,cat1)
hmm9s <- hmm(9,cat1)

hmmnum <- function(s){
  return(s^2 + 2*s -1)
}

hmmsinnum <- function(s){
  return(3*s^2 - 1)
}

moveBIChmm <- function(mm,ns){
  n <- nrow(mm[[1]])
  ll <- mm[[3]][[1]]
  k <- hmmnum(ns)
  
  BIC <- -2*ll + k*log(n)
  return(BIC)
}

movell <- function(mm){
  return(mm[[3]][[1]])
}
fitlist <- list(hmm2s,hmm3s,hmm4s,hmm5s,hmm6s,hmm7s,hmm8s,hmm9s)
library(dplyr)
hmmdf <- data.frame(nstates = 2:9
  , freepars = hmmnum(2:9)
  , ll = -1*unlist(lapply(fitlist,movell))
  , nobs = nrow(cat1)
) %>% mutate(BIC = 2*ll + freepars*log(nobs))

ggplot(hmmdf,aes(x=nstates,y=ll))+geom_point()+geom_line()

```

I did a bunch of BIC stuff too, but the free parameter calcuation is wrong (ie. not sure if I should include the 2n turning angle parameters and n zero-inflation parameters). You can see the problem by looking at the log-likelihood because it is not asymptotic. If you want to play around with it, you can try running states 2 to 5, fitting 7+ will take much more time.

```{r end time,echo=FALSE}
t2 <- proc.time()
t2-t1
```

# Testing time-heterogeneous HMMs
```{r testing time-het}
hmmsin <- function(n,dd){
  mu0 <- c(0:(n-1))
  sd0 <- rep(0.2,n)
  z0 <- rep(0.1,n)
  step0 <- c(mu0,sd0,z0)
  angle0 <- rep(c(0,10000),each=n)
  mod <- fitHMM(data=dd,nbStates=n,stepPar0=step0,anglePar0=NULL,
                stepDist = "lnorm", angleDist = "none",
                formula = ~ cos(2*pi*(as.numeric(time)-1)/24)+ sin(2*pi*(as.numeric(time)-1)/24))
  return(mod)
}

hmmsin2s <- hmmsin(2,cat1)
hmmsin3s <- hmmsin(3,cat1)
hmmsin4s <- hmmsin(4,cat1)

movell(hmm2s)
movell(hmmsin2s)
movell(hmm3s)
movell(hmmsin3s)
movell(hmm4s)
movell(hmmsin4s)
```

I think something is wrong with 2 and 3 state models. Maybe that is the reason why the log-likelihood plot does not look asymptotic.