---
title: "Ordinal Transmation"
author: Michael WZ Li 
date: "`r format(Sys.time(), '%H:%M %d %B %Y')`"
output:
    html_document:
    mathjax: default
---

We want to transform a ordinal fitted object back to the ordinal (cumulative probability) scale? See [here](https://cran.r-project.org/web/packages/ordinal/vignettes/clmm2_tutorial.pdf)

```{r pkg}
library(ordinal)

y <- c(rep(0,100),rep(1,50),rep(2,50),rep(3,50),rep(4,50))
summary(y)
sum(50+100+150+200)/300
set.seed(2830)
y_randomorder <- y[sample(1:length(y),length(y))]
summary(y_randomorder)

factor_y <- factor(y)
dat <- data.frame(y=factor_y,y_ro=factor(y_randomorder))

mod <- clm(y~1,data=dat)
summary(mod)

mod$alpha

```

The cumulative probability is the following:
$$
P(Y \leq j) = \textrm{logit}^{-1}(\alpha_{j})
$$

where $\alpha$'s are the threshold coefficients.

We can compute the above probability by hand:

$$
\textrm{logit}^{-1}(\nu) = \frac{1}{1+exp(\nu)}
$$
, $plogis(-\nu)$ in R.

```{r calculating by hand}
print(1/(1+exp(mod$alpha)))
probs <- plogis(-mod$alpha)
print(probs)
sum(probs)

mod_ro <- clm(y_ro~1,data=dat)

probs_ro <- plogis(-mod_ro$alpha)
print(probs_ro)
sum(probs_ro)

```

## Assignment random grouping for random effect

```{r simple RE}
dat$group <- 1:10
head(dat)
mod_RE <- clmm(y_ro~1+(1|group),data=dat)

summary(mod_RE)
probs_RE <- plogis(-mod_RE$alpha)
print(probs_RE)
sum(probs_RE)

```
As expected, not much change.

### Adding fix effects 
```{r fix}
dat$fix <- 1:2
mod_fix <- clm(y_ro~fix,data=dat)

summary(mod_fix)
probs_fix <- plogis(mod_fix$beta-mod_fix$alpha)
print(probs_fix)
sum(probs_fix)

print(sum(plogis(-mod_fix$alpha)))

```

Did I transform it correctly for the fixed-effect example?
