---
title: "Poisson Epidemic Simulator"
author: Michael Li
date: "`r format(Sys.time(), '%H:%M %d %B %Y')`"
---
  
I am trying to hard code a Poisson process-process and Poisson observation process to fit in JAGS. I want to make sure this "model" can simulate relatively similar data as my chain binomial/ beta-binomial. 

Unlike the binomial family, the Poisson family do not have a population limit and assumes it has an infinite population. Let's create a very simple simulator and assume:

$I_{t} \sim Poisson(I_{t-1})$

Where the mean is the number of infectants of the previous time step (will change later, I don't think it is a stupid idea at this point).

At this point, I am not even going to worry about the observation process because the "process-process" doesn't look right yet. I also do not know how to add FOI / probability of adequate contact per infective. 

Doesn't the poisson automatically handles this? ie: $P(X) = Poisson(\lambda)$ is probability of X new infectants in the new time step. 

```{r pkg,warning=FALSE,message=FALSE}
library(ggplot2)
```

Simulator

```{r simulator,warning=FALSE,message=FALSE}
simm <- function(i0=10,numobs=20,repMean=0.5,shapeP=0.1,seed=NULL){
  if (!is.null(seed)) set.seed(seed)
  tvec <- seq(1,numobs)
  n <- length(tvec)
  I <- Iobs <- IMean <- numeric(n)
  
  I[1] <- i0
#  pShape <- rgamma(1,shape=shapeP,rate=shapeP)
#  IMean[1] <- rgamma(1,shape=pShape,rate=pShape/I[1])
  IMean[1] <- I[1]
  Iobs[1] <- rpois(1,repMean*I[1])
  
  for (t in 2:n){
    I[t] <- rpois(1,IMean[t-1])
#    IMean[t] <- rgamma(1,shape=pShape,rate=pShape/I[t])
    IMean[t] <-I[t]
    Iobs[t] <- rpois(1,repMean*I[t])
  }
  
  data.frame(time=tvec, I, Iobs,IMean)
}
```

Simulate a couple of plots via changing seeds

```{r plots,warning=FALSE,message=FALSE}

p1 <- ggplot(simm(seed=101),aes(x=time,y=I)) + geom_line()

p1
p1 %+% simm(seed=108)
p1 %+% simm(seed=110)

```

