---
title: "Pokemon Experience Curve"
date: "`r format(Sys.time(), '%H:%M %d %B %Y ')`"
author: Michael Li
---

# Exp Curve for Pokemon Go

Pokemon Go raised the Level cap to level 40. It just seems so hard to level up after level 20. I am just curious what the _exp curve_ looks like. There is a *exp table* at  [serbebii.net](http://www.serebii.net/pokemongo/trainer.shtml), but tables are not good. This is a good example illustrating why pictures are better than tables when interpreting data. I should of scrape the table but too lazy.  

```{r dataframe}
lvl <- 1:39
explvl <- (c(1:10,
            10,10,10,15,20,20,20,25,25,50,
            75,100,125,150,190,200,250,300,350,500,
            500,750,1000,1250,1500,2000,2500,3000,5000) * 1000)

dat <- data.frame(Level = lvl,Exp=explvl)

```

# Exp Curve on the Linear Scale 

```{r pkgs}
library(ggplot2)
theme_set(theme_bw())
g <- (ggplot(dat,aes(x=Level,y=Exp))
  + geom_point()
  + geom_smooth()
  + ylab("Experience to Next Level")
)
print(g)

```

# Exp Curve on the Log Scale

```{r log scale}
g+scale_y_log10()
```

##This is insane. I can never reach max level :(
